---
title: "Case Study Code"
author: "Kunwu Lyu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, eval = T, 
                      prompt = FALSE, 
                      cache = TRUE,
                      message = FALSE, warning = FALSE, 
                      fig.height = 3, 
                      # fig.width = 3.5, 
                      fig.align = "center",
                      tidy.opts = list(width.cutoff = 60), tidy = TRUE)
library(tidyverse)
library(lme4)
library(HLMdiag)
library(gridExtra)
library(broom.mixed)
library(knitr)
library(effects)
library(patchwork)
library(mice)
library(GGally)
```

# Modeling

## Data Wrangling

```{r Load in Data}
nurse <- read.csv("https://math.carleton.edu/kstclair/data/bmlr/nursebp.csv",
                  stringsAsFactors = TRUE) %>% 
  as_tibble %>% 
  select(-c(DIA, HRT, timept)) %>% 
  janitor::clean_names() %>% 
  mutate(snum = as.factor(snum))
glimpse(nurse)
out<-md.pattern(nurse, rotate.names=T) #missingness pattern
nurse_complete <- nurse %>% 
  drop_na() %>% 
  mutate(mood = hap - (str + tir)/2,
         standing = case_when(posture == "RECLINE" ~ 0,
                         posture == "SIT" ~ 0,
                         posture == "STAND" ~ 1)) 
# new mood variable

nurse_filtered <- nurse_complete %>% 
  mutate(
    age24 = age-24, 
    phase2 = if_else(phase == "F", 0, 1),
    day2 = if_else(day == "W", 1, 0),
    posture2 = case_when(posture == "RECLINE" ~ 0,
                         posture == "SIT" ~ 1,
                         posture == "STAND" ~ 2),
    fh123 = case_when(fh123 == "NO" ~ 0,
                      fh123 == "YES" ~ 1,
                      fh123 == "YESYES" ~ 2)
         ) %>% 
  mutate(fh123 = as.factor(fh123),
         phase2 = as.factor(phase2),
         day2 = as.factor(day2),
         posture2 = as.factor(posture2)) %>% 
  mutate(day3 = if_else(day == "W", "Workday", "Non-Workday")) %>% 
  mutate(standing = if_else(posture2 == 2, 1, 0),
         fh_yes = if_else(fh123 == 0, 0, 1)) %>% # New standing var
  mutate(standing = as.factor(standing),
         fh_yes = as.factor(fh_yes)) %>% 
  mutate(fh_2 = if_else(fh_yes == 1, "Yes", "No")) %>% 
  drop_na()

nurse_bysubject <- nurse_filtered %>% 
  mutate(phase3 = if_else(phase2 == 0, "Follicular", "Luteal"),
         day3 = if_else(day2 == 1, "Workday", "Non-Workday")) %>% 
  group_by(snum) %>% 
  summarize(
    mean_sys = mean(sys),
    phase3 = first(phase3), day3 = first(day3), 
    age2 = first(age24), fh_2 = first(fh_2), fh123 = first(fh123)
  )

```

## EDA

### Level 1 by Clusters
```{r Level 1 by Clusters}
# Subset of data to viz bp by time
set.seed(93487303)
nurse_all <- nurse_filtered %>% distinct(snum) %>% pull()
random_snums <- sample(nurse_all, 12)
nurse_small <- nurse_filtered %>% filter(snum %in% random_snums)
ggplot(nurse_small, 
       aes(x = timepass, y = sys)) + 
  geom_point() +
  geom_line() + 
  facet_wrap(~snum) +
  labs(x = "Minutes since first measurement",
       y = "BP")

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys)) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "BP by Subjects") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(nurse_filtered, aes(x = timepass, y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Minutes since first measurement",
       y = "BP",
       title = "BP vs. time passed by Subjects",
       color = "Subject ID") +
  guides(color = "none")

ggplot(nurse_filtered, aes(x = mnact5, y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Activity Level",
       y = "BP",
       title = "BP vs. Activity Level",
       color = "Subject ID") +
  guides(color = "none")

ggplot(nurse_filtered, aes(x = as.numeric(standing), y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Standing vs. Not Standing",
       y = "BP",
       title = "BP vs. Posture (2 Levels)",
       color = "Subject ID") +
  guides(color = "none")

ggplot(nurse_filtered, aes(x = mood, y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Mood",
       y = "BP",
       title = "BP vs. Mood",
       color = "Subject ID") +
  guides(color = "none")
```

### Level 2 Covariates
```{r Level 2 Covariates}
## Bivariate
ggplot(nurse_bysubject, aes(x = phase3, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "Menstural Phase",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = day3, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "Workday",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = fh_2, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "With Family History",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = fh123, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "Number of Parents with Hypertension",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = as.factor(age2), y = mean_sys)) +
  geom_boxplot(show.legend = F) + 
  geom_smooth(method = "lm", se=TRUE, aes(group=1)) +
  labs(x = "Age",
       y = "Average Systolic Blood Pressure")

## By age
ggplot(nurse_bysubject, aes(x = age2, y = mean_sys)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ phase3) +
  labs(x = "Age beyond 24",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = age2, y = mean_sys)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ day3) +
  labs(x = "Age beyond 24",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = age2, y = mean_sys)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ fh_2) +
  labs(x = "Age beyond 24 by Family History",
       y = "Average Systolic Blood Pressure")
```

### Level 1 by Level 2 Covariates
```{r Level 1 by Level 2 Covariates}
ggplot(nurse_filtered, aes(x = sys)) +
  geom_histogram(color="white") +
  labs(x = "Systolic Blood Pressure",
       title = "Level 1 BP Distribution")

ggplot(nurse_filtered, aes(x = fh_2, y = sys)) +
  geom_boxplot() + 
  facet_wrap(~ day3) +
  labs(x = "Family History",
       y = "Systolic Blood Pressure",
       title = "Level 1 BP by Workday and Family History")

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys, fill = day3)) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "Level 1 BP by Subjects and Workday",
       fill = "Workday") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = c(0.1, 0.9))

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys, fill = fh_2)) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "Level 1 BP by Subjects and Family History",
       fill = "Family History") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = c(0.1, 0.9))

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys, fill = as.factor(age24))) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "Level 1 BP by Subjects and Age",
       fill = "Age") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = "none")
```

### Spaghetti Plots
```{r Spaghetti Plots}
## Time
ggplot(nurse_filtered, aes(x = timepass, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ day3) +
  labs(x = "Time since first measurement",
       y = "Systolic Blood Pressure")

ggplot(nurse_filtered, aes(x = timepass, y = sys, color = as.factor(day3))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Workday",
       x = "Time since first measurement",
       y = "Systolic Blood Pressure") + theme(legend.position = c(0.2, 0.8))

ggplot(nurse_filtered, aes(x = timepass, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ fh_2)

ggplot(nurse_filtered, aes(x = timepass, y = sys, color = as.factor(fh_2))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Family History",
       x = "Time since first measurement",
       y = "Systolic Blood Pressure") + theme(legend.position = c(0.2, 0.8))


## Activity level
ggplot(nurse_filtered, aes(x = mnact5, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ day3)

ggplot(nurse_filtered, aes(x = mnact5, y = sys, color = as.factor(day3))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Workday",
       x = "Activity Level",
       y = "Systolic Blood Pressure") + theme(legend.position = c(0.2, 0.8))

ggplot(nurse_filtered, aes(x = mnact5, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ fh_2)

ggplot(nurse_filtered, aes(x = mnact5, y = sys, color = as.factor(fh_2))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Family History",
       x = "Activity Level",
       y = "Systolic Blood Pressure") + theme(legend.position = c(0.2, 0.8))


## Posture - 2 levels
ggplot(nurse_filtered, aes(x = standing, y = sys)) +
  geom_boxplot() +
  facet_wrap(~ day3) +
  labs(x = "Standing vs Non Standing",
       y = "Systolic Blood Pressure")

ggplot(nurse_filtered, aes(x = standing, y = sys)) +
  geom_boxplot() +
  facet_wrap(~ fh_2) +
  labs(x = "Standing vs Non Standing",
       y = "Systolic Blood Pressure",
       title = "BP by Standing and Family History") 


## Mood Ratings
ggplot(nurse_filtered, aes(x = mood, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ day3)

ggplot(nurse_filtered, aes(x = mood, y = sys, color = as.factor(day3))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Workday",
       x = "Mood",
       y = "Systolic Blood Pressure") + theme(legend.position = c(0.2, 0.8))

ggplot(nurse_filtered, aes(x = mood, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ fh_2)

ggplot(nurse_filtered, aes(x = mood, y = sys, color = fh_2)) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Family History",
       x = "Mood",
       y = "Systolic Blood Pressure") + theme(legend.position = c(0.2, 0.8))
```

### Separate MLR
```{r Separate MLR}
ls_fits <- group_modify(group_by(nurse_complete, snum), 
                        ~ tidy(lm(sys ~ timepass + mnact5 + standing + mood, 
                                  data =.x))[,1:2])
ls_fits_final <- ls_fits %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename("Intercept"= `(Intercept)`, "Time"= timepass,
         "Activity" = mnact5, "Standing" = standing,
         "Mood" = mood) %>%
  left_join(nurse_bysubject, by = "snum")

## Intercept
ggplot(ls_fits_final, aes(x = day3, y = Intercept)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_final, aes(x = fh_2, y = Intercept)) +
  geom_boxplot() +
  labs(x = "Family History")
ggplot(ls_fits_final, aes(x = age2, y = Intercept)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Age")

## Timepass
ggplot(ls_fits_final, aes(x = day3, y = Time)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_final, aes(x = fh_2, y = Time)) +
  geom_boxplot() +
  labs(x = "Family History")
ggplot(ls_fits_final, aes(x = age2, y = Time)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Age")


## Activity Level
ggplot(ls_fits_final, aes(x = day3, y = Activity)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_final, aes(x = fh_2, y = Activity)) +
  geom_boxplot() +
  labs(x = "Family History")
ggplot(ls_fits_final, aes(x = age2, y = Activity)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Age")


## Standing
ggplot(ls_fits_final, aes(x = day3, y = Standing)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_final, aes(x = fh_2, y = Standing)) +
  geom_boxplot() +
  labs(x = "With Family History")
ggplot(ls_fits_final, aes(x = age2, y = Standing)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Age")


## Mood
ggplot(ls_fits_final, aes(x = day3, y = Mood)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_final, aes(x = fh_2, y = Mood)) +
  geom_boxplot() +
  labs(x = "Number of Parents with Hypertension")
ggplot(ls_fits_final, aes(x = age2, y = Mood)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Age")
```


## Model Selection

### Base Model (Random Intercept)

```{r RI vs. SLR}
nurse_lmm_RI <- lmer(sys ~ timepass + (1 | snum), data = nurse_filtered)
summary(nurse_lmm_RI)

nurse_lm <- lm(sys ~ timepass, data = nurse_filtered)
summary(nurse_lm)


(lrt_RI <- as.numeric(2*(logLik(nurse_lmm_RI, REML = TRUE) - 
                            logLik(nurse_lm, REML = TRUE))))
.5*(1-pchisq(lrt_RI, df = 0))+.5*(1-pchisq(lrt_RI, df = 1))
```


### Random Intercept with Level 1 Covariates

```{r RI with Level 1 Covariates vs. RI}
nurse_lmm_RI_lv1 <- lmer(sys ~ timepass + mnact5 + standing + mood + 
                       (1 | snum), data = nurse_filtered)

summary(nurse_lmm_RI_lv1)
anova(nurse_lmm_RI, nurse_lmm_RI_lv1)

# No mood
nurse_lmm_RI_noMood <- lmer(sys ~ timepass + mnact5 + standing +
                       (1 | snum), data = nurse_filtered)
anova(nurse_lmm_RI_noMood, nurse_lmm_RI_lv1)

# No mood or time
nurse_lmm_RI_noMoodTime <- lmer(sys ~ mnact5 + standing +
                       (1 | snum), data = nurse_filtered) #still use time
anova(nurse_lmm_RI_noMoodTime, nurse_lmm_RI_noMood)
anova(nurse_lmm_RI_noMoodTime, nurse_lmm_RI_lv1)
```


### Random Intercept and Slope

```{r RIS with Stand RE}
nurse_lmm_RIS_stand <- lmer(sys ~ timepass + mnact5 + standing + 
                       (1 + standing | snum), 
                       data = nurse_filtered)
summary(nurse_lmm_RIS_stand)

(lrt_RIS_stand <- as.numeric(2*(logLik(nurse_lmm_RIS_stand, REML = TRUE) - 
                            logLik(nurse_lmm_RI, REML = TRUE))))
.5*(1-pchisq(lrt_RIS_stand, df = 1))+.5*(1-pchisq(lrt_RIS_stand, df = 2))
```

### Random Intercept and Slope with Level 2 Covariates

```{r RIS with Age}
# Age
nurse_lmm_RIS_age <- lmer(sys ~ timepass + mnact5 + standing * age24 + 
                       (1 + standing | snum),
                       data = nurse_filtered)

summary(nurse_lmm_RIS_age)
anova(nurse_lmm_RIS_stand, nurse_lmm_RIS_age)
```

```{r RIS with Workday}
# Workday
nurse_lmm_RIS_day <- lmer(sys ~ day2 * (timepass + standing) + mnact5 + 
                       (1 + standing | snum),
                       data = nurse_filtered)

summary(nurse_lmm_RIS_day)
anova(nurse_lmm_RIS_stand, nurse_lmm_RIS_day)

nurse_lmm_RIS_day_nostand <- lmer(sys ~ day2 * (timepass) + standing + mnact5 + 
                       (1 + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day, nurse_lmm_RIS_day_nostand)
anova(nurse_lmm_RIS_day_nostand, nurse_lmm_RIS_stand)

nurse_lmm_RIS_day_nointer <- lmer(sys ~ day2 + timepass + standing + mnact5 + 
                       (1 + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day_nostand, nurse_lmm_RIS_day_nointer)
```

```{r RIS with Family History and Workday}
# Family history
nurse_lmm_RIS_day_fh <- lmer(sys ~ (day2 + fh_yes) * timepass + 
                                    standing + mnact5 + 
                       (1 + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_day_nostand)
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_stand)
```

```{r Correlation}
nurse_lmm_RIS_day_fh_nocorr <- lmer(sys ~ (day2 + fh_yes) * timepass + 
                                    standing + mnact5 + 
                       (1 + standing || snum),
                       data = nurse_filtered) #Convergence issue, have corr
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_day_fh_nocorr, refit = FALSE)
```


## Model Diagnostics

### Residual Analysis
```{r Load in Residuals}
final_model <- nurse_lmm_RIS_day_fh

nurse_resid1 <- hlm_resid(final_model, level = 1, standardize = TRUE)
nurse_resid2 <- hlm_resid(final_model, level = "snum", include.ls = FALSE)
```

#### Marginal Residuals
```{r Marginal Residuals}
ggplot(data = nurse_resid1, aes(x = .mar.fitted, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Marginal",
       y = "Marginal Residuals") 
ggplot(data = nurse_resid1, aes(x = timepass, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1, aes(x = as.factor(standing), y = .chol.mar.resid)) + 
  geom_boxplot() + 
  # geom_smooth() +
  labs(x = "Standing vs. Not standing",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1, aes(x = mnact5, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Marginal Residuals")
```

#### Conditional Residuals
```{r Conditional (Raw) Residuals}
ggplot(data = nurse_resid1, aes(x = .mar.fitted, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Conditional Means",
       y = "Conditional Residuals") 
ggplot(data = nurse_resid1, aes(x = timepass, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "Conditional Residuals")
ggplot(data = nurse_resid1, aes(x = as.factor(standing), y = .std.resid)) + 
  geom_boxplot() + 
  # geom_smooth() +
  labs(x = "Standing vs. Not standing",
       y = "Conditional Residuals")
ggplot(data = nurse_resid1, aes(x = mnact5, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Conditional Residuals")

ggplot(nurse_resid1, aes(sample = .std.resid)) +
  stat_qq_line()+ 
  stat_qq()
```

#### Try transformations --- Didn't work
```{r Exponential Model}
# Exponential model
nurse_lmm_RIS_logx <- lmer(sys ~ (day2 + fh_yes) * timepass + 
                                    standing + log(mnact5 + 1)  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
nurse_resid1_logx <- hlm_resid(nurse_lmm_RIS_logx, level = 1, standardize = TRUE)

ggplot(data = nurse_resid1_logx, aes(x = .mar.fitted, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Logged Fitted Conditional",
       y = "Conditional Residuals") 
ggplot(data = nurse_resid1_logx, aes(x = `log(mnact5 + 1)`, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Conditional Residuals")
```

```{r Logarithmic Model}
# Logarithmic
nurse_lmm_RIS_logy <- lmer(log(sys) ~ (day2 + fh_yes) * timepass + 
                                    standing + mnact5 + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
nurse_resid1_logy <- hlm_resid(nurse_lmm_RIS_logy, level = 1, standardize = TRUE)

ggplot(data = nurse_resid1_logy, aes(x = .mar.fitted, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Conditional",
       y = "Conditional Residuals") 
ggplot(data = nurse_resid1_logy, aes(x = mnact5, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Conditional Residuals")
```

```{r Power Model}
# power
nurse_lmm_RIS_loglog <- lmer(log(sys) ~ (day2 + fh_yes) * timepass + 
                                    standing + log(mnact5 + 1)  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
nurse_resid1_loglog <- hlm_resid(nurse_lmm_RIS_loglog, level = 1, standardize = TRUE)

ggplot(data = nurse_resid1_loglog, aes(x = .mar.fitted, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Conditional",
       y = "Conditional Residuals") 
ggplot(data = nurse_resid1_loglog, aes(x = `log(mnact5 + 1)`, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Logged Activity Level",
       y = "Conditional Residuals")
```

#### LS Residuals
```{r Least Square Residuals}
ggplot(data = nurse_resid1, aes(x = .mar.fitted, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Values",
       y = "LS Residuals") 
ggplot(data = nurse_resid1, aes(x = timepass, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "LS Residuals")
ggplot(data = nurse_resid1, aes(x = as.factor(standing), y = .std.ls.resid)) + 
  geom_boxplot() + 
  # geom_smooth() +
  labs(x = "Standing vs. Not standing",
       y = "LS Residuals")
ggplot(data = nurse_resid1, aes(x = mnact5, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "LS Residuals")

ggplot(nurse_resid1, aes(sample = .std.ls.resid)) +
  stat_qq_line()+ 
  stat_qq()
ggplot(nurse_resid1, aes(x = as.factor(snum), y = .std.ls.resid)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  labs(x = "Subject ID",
       y = "LS Residuals") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

#### Lv 2 RE
```{r Level 2 Random Effects}
ggplot(nurse_resid2, aes(sample = .ranef.intercept)) +
  stat_qq_line()+ 
  stat_qq()
ggplot(nurse_resid2, aes(sample = .ranef.standing1)) +
  stat_qq_line()+ 
  stat_qq()

nurse_age_phase <- nurse_filtered %>% 
  mutate(snum = as.character(snum)) %>% 
  group_by(snum) %>%
  summarize(nurse_age = first(age24),
            nurse_phase = first(phase)) %>% 
  left_join(nurse_resid2, by = "snum")

## Intercept
ggplot(nurse_age_phase, aes(x = as.factor(nurse_age), y = .ranef.intercept)) +
  geom_boxplot() + 
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se=TRUE, aes(group=1)) +
  labs(title = "RE intercept vs Centered Age",
       x = "Centered Age",
       y = " RE Intercept")
ggplot(nurse_age_phase, aes(x = nurse_phase, y = .ranef.intercept)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  labs(title = "RE intercept vs Menstrual Phase",
       x = "Menstrual Phase",
       y = " RE Intercept")
```


### Influential Statistics

```{r Compute, cache=TRUE}
nurse_infl2 <- hlm_influence(final_model, level = "snum", approx = FALSE, 
                             leverage = c("overall", "fixef", "ranef.uc"))
```

#### Cook's Dist
```{r Cooks Distance}
dotplot_diag(nurse_infl2$cooksd, name = "cooks.distance", cutoff = "internal")

cd_flagged <- nurse_infl2 %>% 
  arrange(desc(cooksd)) %>% 
  slice(1:3) %>% 
  pull(snum)

ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% cd_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2)) +
  labs(x = "Time passed",
       y = "BP")
```

#### RVC
```{r RVC}
dotplot_diag(nurse_infl2$rvc.D11, name = "rvc", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$rvc.D22, name = "rvc", cutoff = "internal", modify = "boxplot")

rvc1_flagged <- nurse_infl2 %>%
  arrange(desc(rvc.D11)) %>%
  slice(1) %>%
  pull(snum)
rvc2_flagged <- nurse_infl2 %>%
  arrange(desc(rvc.D22)) %>%
  slice(1:2) %>%
  pull(snum)

ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% rvc1_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2)) +
  labs(x = "Time passed",
       y = "BP")
  ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% rvc2_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2)) +
  labs(x = "Time passed",
       y = "BP")
```

#### Leverage
```{r Leverage}
dotplot_diag(nurse_infl2$leverage.overall, name = "leverage", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$leverage.fixef, name = "leverage", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$leverage.ranef.uc, name = "leverage", cutoff = "internal", modify = "boxplot")

lev_flagged <- nurse_infl2 %>%
  arrange(desc(leverage.overall)) %>%
  slice(1:3) %>%
  pull(snum)

ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% lev_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2)) +
  labs(x = "Time passed",
       y = "BP")
```


## Effects Interpretation

### Confidence Intervals

```{r Fixed Effects, cache=TRUE}
confint(final_model, parm = "beta_", method = "boot") -> boot_ci
kable(boot_ci)
confint(final_model, parm = "beta_") -> ci
kable(ci)
kable(tidy(final_model))
```

```{r Random Effects}
lattice::dotplot(ranef(final_model))
```


### Effects Plots

```{r Predictor Means}
timepass_all <- nurse_filtered %>% distinct(timepass) %>% pull()
random_timepass <- sample(timepass_all, 3)
# plot(predictorEffects(final_model, xlevels = list(timepass = random_timepass)))
plot(predictorEffects(final_model, xlevels = 4),
                      axes = list(
                        x = list(rug = F, rotate = 90,
                                 day2 = list(lab = "Workday"), 
                                 fh_yes = list(lab = "Family History"), 
                                 timepass = "Time Passed", 
                                 standing = "Standing",
                                 mnact5 = list(lab = "Activity Level"), 
                                 mood = "Mood"),
                        y = list(ticks=list(at=c(110, 114, 118, 122, 126)),
                                 lab = "Blood Pressure"))
                      )
```

```{r Marginal Trend}
nurse_filtered <- nurse_filtered %>%
          # predicted means with RE
  mutate(fit_cond = predict(final_model), 
         fit_marg = predict(final_model, re.form = NA))
          # predicted means without RE

ggplot(nurse_filtered, aes(x=timepass, y = sys)) +
  geom_point() +
  geom_line(aes(y = fit_cond, group = snum), color = "darkgrey") +
  geom_line(aes(y = fit_marg), color = "red", size = 1, alpha = 0.5) +
  facet_grid(rows = vars(day3), cols = vars(fh_2)) + 
  labs(title = "Blood Pressure vs Time Passed, by Workday and Family History",
       x = "Minutes since first measurement",
       y = "Systolic Blood Pressure") +
  theme_minimal(base_size = 17)
```

