---
title: "Case Study Code"
author: "Kunwu Lyu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, eval = T, 
                      prompt = FALSE, 
                      cache = TRUE,
                      message = FALSE, warning = FALSE, 
                      fig.height = 3, 
                      # fig.width = 3.5, 
                      fig.align = "center",
                      tidy.opts = list(width.cutoff = 60), tidy = TRUE)
library(tidyverse)
library(lme4)
library(HLMdiag)
library(gridExtra)
library(broom.mixed)
library(knitr)
library(effects)
library(patchwork)
library(mice)
library(GGally)
```

# Prompt (BLMR)

UCLA nurse blood pressure study. A study by Goldstein and Shapiro (2000) collected information from 203 registered nurses in the Los Angeles area between 24 and 50 years of age on blood pressure (BP) and potential factors that contribute to hypertension. This information includes family history, and whether the subject had one or two hypertensive parents, as well as a wide range of measures of the physical and emotional condition of each nurse throughout the day. **Researchers sought to study the links between BP and family history, personality, mood changes, working status, and menstrual phase**.

Data from this study provided by Weiss (2005) includes observations (40-60 per nurse) repeatedly taken on the 203 nurses over the course of a single day. The first BP measurement was taken half an hour before the subject’s normal start of work, and BP was then measured approximately every 20 minutes for the rest of the day. At each BP reading, the nurses also rate their mood on several dimensions, including how stressed they feel at the moment the BP is taken. In addition, the activity of each subject during the 10 minutes before each reading was measured using an actigraph worn on the waist. Each of the variables in nursebp.csv is described below:

    SNUM: subject identification number
    SYS: systolic blood pressure (mmHg)
    DIA: diastolic blood pressure (mmHg)
    HRT: heart rate (beats per minute)
    MNACT5: activity level (frequency of movements in 1-minute intervals, over a 10-minute period )
    PHASE: menstrual phase (follicular—beginning with the end of menstruation and ending with ovulation, or luteal—beginning with ovulation and ending with pregnancy or menstruation)
    DAY: workday or non-workday
    POSTURE: position during BP measurement—either sitting, standing, or reclining
    STR, HAP, TIR: self-ratings by each nurse of their level of stress, happiness and tiredness at the time of each BP measurement on a 5-point scale, with 5 being the strongest sensation of that feeling and 1 the weakest
    AGE: age in years
    FH123: coded as either NO (no family history of hypertension), YES (1 hypertensive parent), or YESYES (both parents hypertensive)
    time: in minutes from midnight
    timept: number of the measurement that day (approximately 50 for each subject)
    timepass: time in minutes beginning with 0 at time point 1 

Using **systolic blood pressure as the primary response**, write a short report detailing factors that are significantly associated with higher systolic blood pressure. Be sure to support your conclusions with appropriate exploratory plots and multilevel models. In particular, **how are work conditions—activity level, mood, and work status—related to trends in BP levels**? As an appendix to your report, describe your modeling process—how did you arrive at your final model, which covariates are Level One or Level Two, what did you learn from exploratory plots, etc.?

Potential alternative directions: consider diastolic blood pressure or heart rate as the primary response variable, or even try modeling emotion rating using a multilevel model.

# Modeling

## Data Wrangling

```{r Load in Data}
nurse <- read.csv("https://math.carleton.edu/kstclair/data/bmlr/nursebp.csv",
                  stringsAsFactors = TRUE) %>% 
  as_tibble %>% 
  select(-c(DIA, HRT, timept)) %>% 
  janitor::clean_names() %>% 
  mutate(snum = as.factor(snum))
glimpse(nurse)
out<-md.pattern(nurse, rotate.names=T) #missingness pattern
nurse_complete <- nurse %>% 
  drop_na() %>% 
  mutate(mood = hap - (str + tir)/2,
         standing = case_when(posture == "RECLINE" ~ 0,
                         posture == "SIT" ~ 0,
                         posture == "STAND" ~ 1))

nurse_filtered <- nurse_complete %>% 
  mutate(
    age24 = age-24, 
    phase2 = if_else(phase == "F", 0, 1),
    day2 = if_else(day == "W", 1, 0),
    posture2 = case_when(posture == "RECLINE" ~ 0,
                         posture == "SIT" ~ 1,
                         posture == "STAND" ~ 2),
    fh123 = case_when(fh123 == "NO" ~ 0,
                      fh123 == "YES" ~ 1,
                      fh123 == "YESYES" ~ 2)
         ) %>% 
  mutate(fh123 = as.factor(fh123),
         phase2 = as.factor(phase2),
         day2 = as.factor(day2),
         posture2 = as.factor(posture2)) %>% 
  mutate(day3 = if_else(day == "W", "Workday", "Non-Workday")) %>% 
  mutate(standing = if_else(posture2 == 2, 1, 0),
         fh_yes = if_else(fh123 == 0, 0, 1)) %>% 
  mutate(standing = as.factor(standing),
         fh_yes = as.factor(fh_yes)) %>% 
  mutate(fh_2 = if_else(fh_yes == 1, "Yes", "No")) %>% 
  drop_na()

nurse_bysubject <- nurse_filtered %>% 
  mutate(phase3 = if_else(phase2 == 0, "Follicular", "Luteal"),
         day3 = if_else(day2 == 1, "Workday", "Non-Workday")) %>% 
  group_by(snum) %>% 
  summarize(
    mean_sys = mean(sys),
    phase3 = first(phase3), day3 = first(day3), 
    age2 = first(age24), fh_2 = first(fh_2)
  )

```

## EDA

```{r Level 1 by Clusters}
set.seed(93487303)
nurse_all <- nurse_filtered %>% distinct(snum) %>% pull()
random_snums <- sample(nurse_all, 12)
nurse_small <- nurse_filtered %>% filter(snum %in% random_snums)
ggplot(nurse_small, 
       aes(x = timepass, y = sys)) + 
  geom_point() +
  geom_line() + 
  facet_wrap(~snum)

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys)) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "BP by Subjects") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(nurse_filtered, aes(x = timepass, y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Minutes since first measurement",
       y = "BP",
       title = "BP vs. time passed by Subjects",
       color = "Subject ID") +
  guides(color = "none")

ggplot(nurse_filtered, aes(x = mnact5, y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Activity Level",
       y = "BP",
       title = "BP vs. Activity Level",
       color = "Subject ID") +
  guides(color = "none")

ggplot(nurse_filtered, aes(x = as.numeric(standing), y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Standing vs. Not Standing",
       y = "BP",
       title = "BP vs. Posture (2 Levels)",
       color = "Subject ID") +
  guides(color = "none")

ggplot(nurse_filtered, aes(x = mood, y = sys, color = as.factor(snum))) + 
  geom_point(alpha = 0.3) +
  geom_line(stat = "smooth", method = "lm", se = F, alpha = 0.8) +
    labs(x = "Mood",
       y = "BP",
       title = "BP vs. Mood",
       color = "Subject ID") +
  guides(color = "none")
```

```{r Level 1 by Level 2 Covariates}
ggplot(nurse_filtered, aes(x = sys)) +
  geom_histogram(color="white") +
  labs(x = "Systolic Blood Pressure",
       title = "Level 1 BP Distribution")

ggplot(nurse_filtered, aes(x = fh_2, y = sys)) +
  geom_boxplot() + 
  facet_wrap(~ day3) +
  labs(x = "Family History",
       y = "Systolic Blood Pressure",
       title = "Level 1 BP by Workday and Family History")

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys, fill = day3)) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "Level 1 BP by Subjects and Workday",
       fill = "Workday") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys, fill = fh_2)) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "Level 1 BP by Subjects and Workday",
       fill = "Family History") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(nurse_filtered, aes(x = as.factor(snum), y = sys, fill = as.factor(age24))) +
  geom_boxplot() +
  labs(x = "Subject ID",
       y = "BP",
       title = "Level 1 BP by Subjects and Workday",
       fill = "Family History") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(fill = "none")
```

```{r Level 2 Covariates}
## Bivariate
ggplot(nurse_bysubject, aes(x = phase3, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "Menstural Phase",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = day3, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "Workday",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = fh_2, y = mean_sys)) +
  geom_boxplot() + 
  labs(x = "With Family History",
       y = "Average Systolic Blood Pressure")

## By age
ggplot(nurse_bysubject, aes(x = age2, y = mean_sys)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ phase3) +
  labs(x = "Age beyond 24",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = age2, y = mean_sys)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ day3) +
  labs(x = "Age beyond 24",
       y = "Average Systolic Blood Pressure")

ggplot(nurse_bysubject, aes(x = age2, y = mean_sys)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ fh_2) +
  labs(x = "Age beyond 24 by Family History",
       y = "Average Systolic Blood Pressure")
```

```{r Spaghetti Plots}
## Time
ggplot(nurse_filtered, aes(x = timepass, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ day3) +
  labs(x = "Time since first measurement",
       y = "Systolic Blood Pressure")

ggplot(nurse_filtered, aes(x = timepass, y = sys, color = as.factor(day3))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Workday",
       x = "Time since first measurement",
       y = "Systolic Blood Pressure")

ggplot(nurse_filtered, aes(x = timepass, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ fh_2)

ggplot(nurse_filtered, aes(x = timepass, y = sys, color = as.factor(fh_2))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Family History",
       x = "Time since first measurement",
       y = "Systolic Blood Pressure")


## Activity level
ggplot(nurse_filtered, aes(x = mnact5, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ day)

ggplot(nurse_filtered, aes(x = mnact5, y = sys, color = as.factor(day3))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Workday",
       x = "Activity Level",
       y = "Systolic Blood Pressure")

ggplot(nurse_filtered, aes(x = mnact5, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ fh_2)

ggplot(nurse_filtered, aes(x = mnact5, y = sys, color = as.factor(fh_2))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Family History",
       x = "Activity Level",
       y = "Systolic Blood Pressure")


## Posture - 2 levels
ggplot(nurse_filtered, aes(x = standing, y = sys)) +
  geom_boxplot() +
  facet_wrap(~ day3) +
  labs(x = "Standing vs Non Standing",
       y = "Systolic Blood Pressure")

ggplot(nurse_filtered, aes(x = standing, y = sys)) +
  geom_boxplot() +
  facet_wrap(~ fh_2) +
  labs(x = "Standing vs Non Standing",
       y = "Systolic Blood Pressure",
       title = "BP by Standing and Family History")


## Mood Ratings
ggplot(nurse_filtered, aes(x = mood, y = sys)) +
  geom_line(aes(group = snum), color = "dark grey") +
  geom_smooth(method = "loess", color = "black", se = F, size = 1) +
  facet_wrap(~ day)

ggplot(nurse_filtered, aes(x = mood, y = sys, color = as.factor(day3))) +
  geom_line(aes(group = interaction(snum, day)), alpha = 0.1) +
  geom_smooth(method = "loess", se = FALSE, size = 1) +
  labs(color = "Workday",
       x = "Mood",
       y = "Systolic Blood Pressure")
```

```{r Separate LS}

ls_fits_timepass <- group_modify(group_by(nurse_complete, snum), 
                        ~ tidy(lm(sys ~ timepass, data =.x))[,1:2])
ls_fits_timepass_final <- ls_fits_timepass %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename("Intercept"= `(Intercept)`, "Slope"= timepass) %>%
  left_join(nurse_bysubject, by = "snum")

## Intercept
ggplot(ls_fits_timepass_final, aes(x = day3, y = Intercept)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_timepass_final, aes(x = fh_2, y = Intercept)) +
  geom_boxplot() +
  labs(x = "Family History")
ggplot(ls_fits_timepass_final, aes(x = age2, y = Intercept)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)

## Timepass
ggplot(ls_fits_timepass_final, aes(x = day3, y = Slope)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_timepass_final, aes(x = fh_2, y = Slope)) +
  geom_boxplot() +
  labs(x = "Family History")
ggplot(ls_fits_timepass_final, aes(x = age2, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)


## Activity Level
ls_fits_act <- group_modify(group_by(nurse_complete, snum), 
                        ~ tidy(lm(sys ~ mnact5, data =.x))[,1:2])
ls_fits_act_final <- ls_fits_act %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename("Intercept"= `(Intercept)`, "Slope"= mnact5) %>%
  left_join(nurse_bysubject, by = "snum")

ggplot(ls_fits_act_final, aes(x = day3, y = Slope)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_act_final, aes(x = fh_2, y = Slope)) +
  geom_boxplot() +
  labs(x = "Family History")
ggplot(ls_fits_act_final, aes(x = age2, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)


## Posture
ls_fits_post <- group_modify(group_by(nurse_complete, snum), 
                        ~ tidy(lm(sys ~ standing, data =.x))[,1:2])
ls_fits_post_final <- ls_fits_post %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename("Nonstanding"= `(Intercept)`, "Standing"= standing) %>%
  left_join(nurse_bysubject, by = "snum")
### Non-standing
ggplot(ls_fits_post_final, aes(x = day3, y = Nonstanding)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_post_final, aes(x = fh_2, y = Nonstanding)) +
  geom_boxplot() +
  labs(x = "With Family History")
ggplot(ls_fits_post_final, aes(x = age2, y = Nonstanding)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
### Stand
ggplot(ls_fits_post_final, aes(x = day3, y = Standing)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_post_final, aes(x = fh_2, y = Standing)) +
  geom_boxplot() +
  labs(x = "With Family History")
ggplot(ls_fits_post_final, aes(x = age2, y = Standing)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)


## Mood
ls_fits_mood <- group_modify(group_by(nurse_complete, snum), 
                        ~ tidy(lm(sys ~ mood, data =.x))[,1:2])
ls_fits_mood_final <- ls_fits_mood %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  rename("Intercept"= `(Intercept)`, "Slope"= mood) %>%
  left_join(nurse_bysubject, by = "snum")

ggplot(ls_fits_mood_final, aes(x = day3, y = Slope)) +
  geom_boxplot() +
  labs(x = "Workday")
ggplot(ls_fits_mood_final, aes(x = fh_2, y = Slope)) +
  geom_boxplot() +
  labs(x = "Number of Parents with Hypertension")
ggplot(ls_fits_mood_final, aes(x = age2, y = Slope)) +
  geom_point() +
  geom_smooth(method = "lm", se = F)
```


## Model Selection

### Base Model (Random Intercept)

```{r RI vs. SLR}
nurse_lmm_RI <- lmer(sys ~ timepass + (1 | snum), data = nurse_filtered)
summary(nurse_lmm_RI)

nurse_lm <- lm(sys ~ timepass, data = nurse_filtered)
summary(nurse_lm)


(lrt_RI <- as.numeric(2*(logLik(nurse_lmm_RI, REML = TRUE) - 
                            logLik(nurse_lm, REML = TRUE))))
.5*(1-pchisq(lrt_RI, df = 0))+.5*(1-pchisq(lrt_RI, df = 1))
```


### Random Intercept with Level 1 Covariates

```{r RI with Level 1 Covariates vs. RI}
nurse_lmm_RI_lv1 <- lmer(sys ~ timepass + mnact5 + standing + mood + 
                       (1 | snum), data = nurse_filtered)

summary(nurse_lmm_RI_lv1)
anova(nurse_lmm_RI, nurse_lmm_RI_lv1)

nurse_lmm_RI_noMood <- lmer(sys ~ timepass + mnact5 + standing +
                       (1 | snum), data = nurse_filtered)
anova(nurse_lmm_RI_noMood, nurse_lmm_RI_lv1)

nurse_lmm_RI_noMoodTime <- lmer(sys ~ mnact5 + standing +
                       (1 | snum), data = nurse_filtered)
anova(nurse_lmm_RI_noMoodTime, nurse_lmm_RI_noMood)
anova(nurse_lmm_RI_noMoodTime, nurse_lmm_RI_lv1)
```


### Random Intercept and Slope

```{r RIS with Mood RE}
nurse_lmm_RIS_mood <- lmer(sys ~ timepass + mnact5 + standing + mood + 
                       (1 + mood | snum), 
                       data = nurse_filtered)
summary(nurse_lmm_RIS_mood)

(lrt_RIS_mood <- as.numeric(2*(logLik(nurse_lmm_RIS_mood, REML = TRUE) - 
                            logLik(nurse_lmm_RI, REML = TRUE))))
.5*(1-pchisq(lrt_RI, df = 1))+.5*(1-pchisq(lrt_RI, df = 2))
```

```{r RIS with Mood and Stand RE}
nurse_lmm_RIS_mood_stand <- lmer(sys ~ timepass + mnact5 + standing + mood + 
                       (1 + mood + standing | snum), 
                       data = nurse_filtered)
summary(nurse_lmm_RIS_mood_stand)

(lrt_RIS_mood_stand <- as.numeric(2*(logLik(nurse_lmm_RIS_mood_stand, REML = TRUE) - 
                            logLik(nurse_lmm_RIS_mood, REML = TRUE))))
.5*(1-pchisq(lrt_RIS_mood_stand, df = 2))+.5*(1-pchisq(lrt_RIS_mood_stand, df = 3))
```


### Random Intercept and Slope with Level 2 Covariates

```{r RIS with Age}
nurse_lmm_RIS_age <- lmer(sys ~ timepass + mnact5 + standing + mood + 
                             age24 + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)

summary(nurse_lmm_RIS_age)
anova(nurse_lmm_RIS_mood_stand, nurse_lmm_RIS_age)
```

```{r RIS with Workday}
nurse_lmm_RIS_day <- lmer(sys ~ day2 * (timepass + standing) + mnact5  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)

summary(nurse_lmm_RIS_day)
anova(nurse_lmm_RIS_mood_stand, nurse_lmm_RIS_day)

nurse_lmm_RIS_day_nostand <- lmer(sys ~ day2 * (timepass) + standing + mnact5  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day, nurse_lmm_RIS_day_nostand)
anova(nurse_lmm_RIS_day_nostand, nurse_lmm_RIS_mood_stand)

nurse_lmm_RIS_day_nointer <- lmer(sys ~ day2 + timepass + standing + mnact5  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day_nostand, nurse_lmm_RIS_day_nointer)
```

```{r RIS with Family History and Workday}
nurse_lmm_RIS_day_fh <- lmer(sys ~ (day2 + fh_yes) * timepass + 
                                    standing + mnact5  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_day_nostand)
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_mood_stand)

nurse_lmm_RIS_day_fh_noint <- lmer(sys ~ day2 * timepass + fh_yes +
                                    standing + mnact5  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_day_fh_noint)
```

```{r Correlation}
nurse_lmm_RIS_day_fh_nocorr <- lmer(sys ~ (day2 + fh_yes) * timepass + 
                                    standing + mnact5  + mood + 
                       (1 + mood + standing || snum),
                       data = nurse_filtered)
anova(nurse_lmm_RIS_day_fh, nurse_lmm_RIS_day_fh_nocorr, refit = FALSE)
```


## Model Diagnostics

### Residual Analysis
```{r Load in Residuals}
final_model <- nurse_lmm_RIS_day_fh

nurse_resid1 <- hlm_resid(final_model, level = 1, standardize = TRUE)
nurse_resid2 <- hlm_resid(final_model, level = "snum", include.ls = FALSE)
```

```{r Marginal Residuals}
ggplot(data = nurse_resid1, aes(x = .mar.fitted, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Marginal",
       y = "Marginal Residuals") 
ggplot(data = nurse_resid1, aes(x = timepass, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1, aes(x = as.factor(standing), y = .chol.mar.resid)) + 
  geom_boxplot() + 
  # geom_smooth() +
  labs(x = "Standing vs. Not standing",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1, aes(x = mnact5, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1, aes(x = mood, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Mood",
       y = "Marginal Residuals")
```

```{r Exponential Model}
nurse_lmm_RIS_logx <- lmer(sys ~ (day2 + fh_yes) * timepass + 
                                    standing + log(mnact5 + 1)  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
nurse_resid1_log <- hlm_resid(nurse_lmm_RIS_logx, level = 1, standardize = TRUE)

ggplot(data = nurse_resid1_log, aes(x = .mar.fitted, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Logged Fitted Marginal",
       y = "Marginal Residuals") 
ggplot(data = nurse_resid1_log, aes(x = timepass, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1_log, aes(x = as.factor(standing), y = .chol.mar.resid)) + 
  geom_boxplot() + 
  labs(x = "Standing vs. Not standing",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1_log, aes(x = `log(mnact5 + 1)`, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1_log, aes(x = mood, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Mood",
       y = "Marginal Residuals")
```

```{r Power Model}
nurse_lmm_RIS_loglog <- lmer(log(sys) ~ (day2 + fh_yes) * timepass + 
                                    standing + log(mnact5 + 1)  + mood + 
                       (1 + mood + standing | snum),
                       data = nurse_filtered)
nurse_resid1_loglog <- hlm_resid(nurse_lmm_RIS_loglog, level = 1, standardize = TRUE)

ggplot(data = nurse_resid1_loglog, aes(x = .mar.fitted, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Marginal",
       y = "Marginal Residuals") 
ggplot(data = nurse_resid1_loglog, aes(x = timepass, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1_loglog, aes(x = as.factor(standing), y = .chol.mar.resid)) + 
  geom_boxplot() + 
  labs(x = "Standing vs. Not standing",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1_loglog, aes(x = `log(mnact5 + 1)`, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Logged Activity Level",
       y = "Marginal Residuals")
ggplot(data = nurse_resid1_loglog, aes(x = mood, y = .chol.mar.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Mood",
       y = "Marginal Residuals")
```

```{r Conditional (Raw) Residuals}
ggplot(data = nurse_resid1, aes(x = .mar.fitted, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Conditional Means",
       y = "Conditional Residuals") 
ggplot(data = nurse_resid1, aes(x = timepass, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "Conditional Residuals")
ggplot(data = nurse_resid1, aes(x = as.factor(standing), y = .std.resid)) + 
  geom_boxplot() + 
  # geom_smooth() +
  labs(x = "Standing vs. Not standing",
       y = "Conditional Residuals")
ggplot(data = nurse_resid1, aes(x = mnact5, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "Conditional Residuals")
ggplot(data = nurse_resid1, aes(x = mood, y = .std.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Mood",
       y = "Conditional Residuals")

ggplot(nurse_resid1, aes(sample = .std.resid)) +
  stat_qq_line()+ 
  stat_qq()
```

```{r Least Square Residuals}
ggplot(data = nurse_resid1, aes(x = .mar.fitted, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Fitted Values",
       y = "LS Residuals") 
ggplot(data = nurse_resid1, aes(x = timepass, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Minutes since first measurement",
       y = "LS Residuals")
ggplot(data = nurse_resid1, aes(x = as.factor(standing), y = .std.ls.resid)) + 
  geom_boxplot() + 
  # geom_smooth() +
  labs(x = "Standing vs. Not standing",
       y = "LS Residuals")
ggplot(data = nurse_resid1, aes(x = mnact5, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Activity Level",
       y = "LS Residuals")
ggplot(data = nurse_resid1, aes(x = mood, y = .std.ls.resid)) + 
  geom_point(alpha=0.2) + 
  geom_smooth() +
  labs(x = "Mood",
       y = "LS Residuals")

ggplot(nurse_resid1, aes(sample = .std.ls.resid)) +
  stat_qq_line()+ 
  stat_qq()
ggplot(nurse_resid1, aes(x = as.factor(snum), y = .std.ls.resid)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  labs(x = "Subject ID",
       y = "LS Residuals") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r Level 2 Random Effects}
ggplot(nurse_resid2, aes(sample = .ranef.intercept)) +
  stat_qq_line()+ 
  stat_qq()
ggplot(nurse_resid2, aes(sample = .ranef.mood)) +
  stat_qq_line()+ 
  stat_qq()
ggplot(nurse_resid2, aes(sample = .ranef.standing1)) +
  stat_qq_line()+ 
  stat_qq()

nurse_age_phase <- nurse_filtered %>% 
  mutate(snum = as.character(snum)) %>% 
  group_by(snum) %>%
  summarize(nurse_age = first(age24),
            nurse_phase = first(phase)) %>% 
  left_join(nurse_resid2, by = "snum")

## Intercept
ggplot(nurse_age_phase, aes(x = as.factor(nurse_age), y = .ranef.intercept)) +
  geom_boxplot() + 
  geom_hline(yintercept = 0) +
  geom_smooth(method = "lm", se=TRUE, aes(group=1)) +
  labs(title = "RE intercept vs Centered Age",
       x = "Centered Age",
       y = " RE Intercept")
ggplot(nurse_age_phase, aes(x = nurse_phase, y = .ranef.intercept)) +
  geom_boxplot() +
  geom_hline(yintercept = 0) +
  labs(title = "RE intercept vs Centered Age",
       x = "Menstrual Phase",
       y = " RE Intercept")

```


### Influential Statistics

```{r Compute, cache=TRUE}
nurse_infl2 <- hlm_influence(final_model, level = "snum", approx = FALSE, 
                             leverage = c("overall", "fixef", "ranef.uc"))
```

```{r Cooks Distance}
dotplot_diag(nurse_infl2$cooksd, name = "cooks.distance", cutoff = "internal")

cd_flagged <- nurse_infl2 %>% 
  arrange(desc(cooksd)) %>% 
  slice(1:3) %>% 
  pull(snum)

ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% cd_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2))
```

```{r RVC}
dotplot_diag(nurse_infl2$rvc.D11, name = "rvc", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$rvc.D22, name = "rvc", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$rvc.D33, name = "rvc", cutoff = "internal", modify = "boxplot")

# rvc1_flagged <- nurse_infl2 %>% 
#   arrange(rvc.D11) %>% 
#   slice(1:3)
# rvc2_flagged <- nurse_infl2 %>% 
#   arrange(rvc.D22) %>% 
#   slice(1) %>% 
#   full_join(rvc1_flagged)
# rvc3_flagged <- nurse_infl2 %>% 
#   arrange(rvc.D33) %>% 
#   slice(1:2) %>% 
#   full_join(rvc2_flagged) %>% 
#   pull(snum)

rvc1_flagged <- nurse_infl2 %>%
  arrange(rvc.D11) %>%
  slice(1:3) %>%
  pull(snum)
rvc2_flagged <- nurse_infl2 %>%
  arrange(rvc.D22) %>%
  slice(1) %>%
  pull(snum)
rvc3_flagged <- nurse_infl2 %>%
  arrange(rvc.D33) %>%
  slice(1:2) %>%
  pull(snum)

ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% rvc1_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2))
ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% rvc2_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2))
ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum %in% rvc3_flagged), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2))
```

```{r Leverage}
dotplot_diag(nurse_infl2$leverage.overall, name = "leverage", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$leverage.fixef, name = "leverage", cutoff = "internal", modify = "boxplot")
dotplot_diag(nurse_infl2$leverage.ranef.uc, name = "leverage", cutoff = "internal", modify = "boxplot")

ggplot(nurse_filtered, aes(x = timepass, y = sys))+
  geom_point() + 
  geom_point(data = nurse_filtered %>% filter(snum == 1062 | snum == 1226), color = "red") +
  geom_smooth(method = "lm", se = F) +
  facet_grid(rows = vars(day3), cols = vars(fh_2))
```


## Effects Interpretation

### Confidence Intervals

```{r Fixed Effects, cache=TRUE}
confint(final_model, parm = "beta_", method = "boot")
confint(final_model, parm = "beta_")
```

```{r Random Effects}
lattice::dotplot(ranef(final_model))
```


### Effects Plots

```{r Predictor Means}
timepass_all <- nurse_filtered %>% distinct(timepass) %>% pull()
random_timepass <- sample(timepass_all, 3)
# plot(predictorEffects(final_model, xlevels = list(timepass = random_timepass)))
plot(predictorEffects(final_model, xlevels = 4),
                      axes = list(
                        x = list(rug = F, rotate = 90,
                                 day2 = list(lab = "Workday"), 
                                 fh_yes = list(lab = "Family History"), 
                                 timepass = "Time Passed", 
                                 standing = "Standing",
                                 mnact5 = list(lab = "Activity Level"), 
                                 mood = "Mood"),
                        y = list(ticks=list(at=c(110, 114, 118, 122, 126)),
                                 lab = "Blood Pressure"))
                      )
```

```{r Marginal Trend}
nurse_filtered <- nurse_filtered %>%
          # predicted means with RE
  mutate(fit_cond = predict(final_model), 
         fit_marg = predict(final_model, re.form = NA))
          # predicted means without RE

ggplot(nurse_filtered, aes(x=timepass, y = sys)) +
  geom_point() +
  geom_line(aes(y = fit_cond, group = snum), color = "darkgrey") +
  geom_line(aes(y = fit_marg), color = "red", size = 1, alpha = 0.5) +
  facet_grid(rows = vars(day3), cols = vars(fh_2)) + 
  labs(title = "Blood Pressure vs Time Passed, by Workday and Family History",
       x = "Minutes since first measurement",
       y = "Systolic Blood Pressure") +
  theme_minimal(base_size = 17)
```

